---
title: "Aides Drug"
author: "Sarah Medina"
date: "09/22/2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction
Using the data from the triumpsal clinical trials I will try to determine if the drug appears to be effective or not. We will only consider the results of the drug users to determine if the drug is effective or not. We will obtain this subpopulation using dplyr and the filer command.

```{r eval=FALSE}
triUmp <- triUmp %>%
mutate(across(where(is.character), as.factor))
```
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods
We will be using two helper files Mytricks and Cat_Cat to solve this problem. The Mytricks helper file will allow me to change the character variables into factor variables. We will use Numerical, Graphical, and Inferential methods. In order to obtain the subpopulation of the drug users we will use dplyr with the filter command. 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

```{r}
DrugUser<-filter(triUmp, DrugUser=="yes" )
```
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The Graph demonstrates that the drug is effective. The percentage of the people that got aids and used the drug is significantly lower than the ones who got aids and took the placebo. Of those that got aids 25 percent had taken the drug and 75 percent had taken the placebo. This indicates the drug is effective because the percentages are significantly different from one another. 

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

The numerical results indicate that of the people who were infected with aids 24.29 percent used the drug and got sick. It also demonstrates that of the people who got infected with aids 75.71 percent took the placebo and got sick. This indicates the drug is effective. 
## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher's Exact Test indicates that the odds ratio for a person who took the drug and their likelihood that they will get aids is 0.3534128. Therefore, you are three times less likely to get infected with aids if you take the drug and three times more likely to get infected with aids if you don't take the drug. These results are with respect to the subpopulation of the DrugUsers.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

After analyzing the sub population of the DrugUsers. The results indicate that the DrugUsers are three times more likely to get infected with aids if they do not take the drug. DrugUsers are also three times less likely to get infected with aids if they take the drug. Therefore, it is recommended that the sub population of DrugUsers take the drug in order to prevent being infected with aids. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
